Unshuffling Permutations

### 2015-09-26 submission.tex
Typos

  It is immediate that H can be constructed in polynomial-time in $n$ and $k$.
  % We claim that $\sigma$ occurs in $\pi$ if and only if
  % there exists a containment-free perfect matching
  % $\mathcal{M}$ in $G_\mu$ such that
  % for any two disctinct edges
  % $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  % we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
  It can be shown (see Appendix) that
  $\sigma$ occurs in $\pi$ if and only if
  there exists a containment-free perfect matching
  $\mathcal{M}$ in $K_\mu$ such that
  for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
  \qed
\end{proof}

@@

\begin{proof}[of Claim~\ref{claim:binary_to_permutation_avoiding}]
    Let us first show that the image of $\BINTOPERM$ contains only
    permutations avoiding $213$ and $231$. Let $u$ be a binary word,
    $\pi = \BINTOPERM(u)$, and $P_0$ (resp. $P_1$) be the set of the
    positions of the occurrences of $0$ (resp. $1$) in $u$. By
    definition of $\BINTOPERM$, from left to right, the subword
    $v = \pi_{|P_0}$ is increasing and the subword $w = \pi_{|P_1}$
    is decreasing, and all letters of $w$ are greater than those
    of~$v$. Now, assume that $\pi$ admits an occurrence of $213$.
    Then, since $v$ is increasing and $w$ is decreasing, there is an
    occurrence of $3$ (resp. $13$, $23$) in $v$ and a relative
    occurrence of $21$ (resp. $2$, $1$). All these three cases
    contradict the fact that all letters of $w$ are greater than
    those of $v$. A similar argument shows that $\pi$ avoids~$231$
    as well.
    \smallskip

@@

    (\emph{1} $\Rightarrow$ \emph{2}).
    Since $\pi$ is a square, it admits a square root $\sigma$. Then, by
    definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
    and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
    letters of $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
    order-isomorphic to $\sigma$. Let $2n = |\pi|$ (and hence
    $|\sigma| = n$). Let $G_\pi$ be the linear clique associated to $\pi$
    with vertices
    $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
    is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
    is a subset of $\ES{G_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
    matching since $I^1 \cap I^2 = \emptyset$, and
    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$. It is also containment-free.
    Indeed, if it is not the case then there would exist two edges
    $e = \{i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$, $j < k$, in
    $\mathcal{M}$ such that $i^1_j < i^1_k < i^2_k < i^2_j$ or
    $i^1_j < i^2_k < i^1_k < i^2_j$. This is a contradiction since
    $i^2_k > i^2_j$ if $k > j$. Finally, for any two distinct edges
    $e = \{\i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
    we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
    $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing in both cases two
    elements (at position $j$ and $k$) in two patterns that are
    order-isomorphic to $\sigma$.

@@

  Conversely, suppose that there exists a containment-free perfect matching
  $\mathcal{M}$ in $G_\mu$ such that for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
  We show that $\sigma$ occurs as a pattern in $\pi$.
  Whereas the matching $\mathcal{M}$ may not be as regular as in the forward
  direction, the main idea is to prove that $\mathcal{M}$ contains enough structure
  (more precisely, $k+2$ $(\sigma', \pi')$-edges) so that we can conclude
  that $\sigma$ occurs in $\pi$.
  We have divided the reverse direction into a serie of basic claims that
  progressively defines the overall structure of $\mathcal{M}$.

>>>>

  It is immediate that H can be constructed in polynomial-time in $n$ and $k$.
  % We claim that $\sigma$ occurs in $\pi$ if and only if
  % there exists a containment-free perfect matching
  % $\mathcal{M}$ in $K_\mu$ such that
  % for any two disctinct edges
  % $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  % we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
  It can be shown (see Appendix) that
  $\sigma$ occurs in $\pi$ if and only if
  there exists a containment-free perfect matching
  $\mathcal{M}$ in $K_\mu$ such that
  for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
  \qed
\end{proof}

@@

\begin{proof}[of Claim~\ref{claim:binary_to_permutation_avoiding}]
    Let us first show that the image of $\BINTOPERM$ contains only
    permutations avoiding $213$ and $231$. Let $u$ be a binary word,
    $\pi = \BINTOPERM(u)$, and $P_0$ (resp. $P_1$) be the set of the
    positions of the occurrences of $0$ (resp. $1$) in $u$. By
    definition of $\BINTOPERM$, from left to right, the subword
    $v = \pi_{|P_0}$ is increasing and the subword $w = \pi_{|P_1}$
    is decreasing, and all letters of $w$ are greater than those
    of~$v$. Now, assume that $\pi$ admits an occurrence of $213$.
    Then, since $v$ is increasing and $w$ is decreasing, there is an
    occurrence of $3$ (resp. $13$, $23$) in $v$ and a relative
    occurrence of $21$ (resp. $2$, $1$) in $w$. All these three cases
    contradict the fact that all letters of $w$ are greater than
    those of $v$. A similar argument shows that $\pi$ avoids~$231$
    as well.
    \smallskip

@@

    (\emph{1} $\Rightarrow$ \emph{2}).
    Since $\pi$ is a square, it admits a square root $\sigma$. Then, by
    definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
    and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
    letters of $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
    order-isomorphic to $\sigma$. Let $2n = |\pi|$ (and hence
    $|\sigma| = n$). Let $K_\pi$ be the linear clique associated to $\pi$
    with vertices
    $\VS{K_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
    is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
    is a subset of $\ES{K_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
    matching since $I^1 \cap I^2 = \emptyset$, and
    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$. It is also containment-free.
    Indeed, if it is not the case then there would exist two edges
    $e = \{i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$, $j < k$, in
    $\mathcal{M}$ such that $i^1_j < i^1_k < i^2_k < i^2_j$ or
    $i^1_j < i^2_k < i^1_k < i^2_j$. This is a contradiction since
    $i^2_k > i^2_j$ if $k > j$. Finally, for any two distinct edges
    $e = \{\i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
    we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
    $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing in both cases two
    elements (at position $j$ and $k$) in two patterns that are
    order-isomorphic to $\sigma$.

@@

  Conversely, suppose that there exists a containment-free perfect matching
  $\mathcal{M}$ in $K_\mu$ such that for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
  We show that $\sigma$ occurs as a pattern in $\pi$.
  Whereas the matching $\mathcal{M}$ may not be as regular as in the forward
  direction, the main idea is to prove that $\mathcal{M}$ contains enough structure
  (more precisely, $k+2$ $(\sigma', \pi')$-edges) so that we can conclude
  that $\sigma$ occurs in $\pi$.
  We have divided the reverse direction into a serie of basic claims that
  progressively defines the overall structure of $\mathcal{M}$.
